---
title: "Reproducible Research: Peer Assessment 1"
---


## Loading and preprocessing the data 
```{r}
unzip('activity.zip')
activity <- read.csv('activity.csv', sep=",",quote="\"", dec=".", fill=TRUE)
````

## What is mean total number of steps taken per day?
- Remove incomplete observations
```{r}
activity_clear <- activity[complete.cases(activity),]
activity_clear$date <- as.Date(activity_clear$date) 
```
- load *plyr* library
```{r}
library(plyr)
```
- create new data frame with sum of steps taken each day
```{r}
StepsDaily <- ddply(activity_clear,~date,summarise,sum=sum(steps))
```
- plot histogram
```{r}
library(ggplot2)
print( qplot(StepsDaily$date,StepsDaily$sum, geom='bar', stat='identity') )
summary(StepsDaily)
```

- calulate mean
```{r}
mean(StepsDaily$sum)
```
- calulate median
```{r}
median(StepsDaily$sum)
```
## What is the average daily activity pattern?
- create necessary data frame
```{r}
StepsInterval <- ddply(activity_clear,~interval,summarize,mean=mean(steps))
```
- plot a line that show mean number of steps taken at given interval
```{r}
plot(StepsInterval$mean, type="l")
```
- identify interval with maximum average number of steps
```{r}
StepsInterval[max(StepsInterval$mean),]
```
## Inputing missing values
- identify subset that has missing values
```{r}
activity_missing <- activity[is.na(activity),]
```
- assign zero to all missing values
```{r}
activity_missing$steps <- 0
```
- creates a new data frame that merges data from missing subset with average data
```{r}
activity_missing_completed <- merge(activity_missing,StepsInterval, by.x = "interval", by.y = "interval")
```
- assing a mean to all missing values and drop unnecessary column
```{r}
activity_missing_completed$steps <- activity_missing_completed$mean
activity_missing_completed$mean <- NULL
```
- binds both complete.cases dataframes 
```{r}
activity_na_imput <- rbind(activity_clear,activity_missing_completed)
```
- creates sums over particular dates (with new data) and plots appropriate chart
```{r}
StepsDaily_na_imput <- ddply(activity_na_imput,~date,summarise,sum=sum(steps))
print( qplot(StepsDaily_na_imput$date,StepsDaily_na_imput$sum, geom='bar', stat='identity') )
```
- new mean value
```{r}
mean(StepsDaily_na_imput$sum)
```
- new median value
```{r}
median(StepsDaily_na_imput$sum)
```
## Are there differences in activity patterns between weekdays and weekends?
- adds another column that changes the date into weekday name
```{r}
activity_na_imput$day <- weekdays(as.Date(activity_na_imput$date))
```
- creates a data frame that converts day name to weekeday or weekend, accordingly
```{r}
weekend_frame <- data.frame(day=c("Monday","Tuesday", "Wednesday", "Thursday","Friday","Saturday","Sunday"),weekend=c("weekday","weekday","weekday","weekday","weekday","weekend","weekend"))
```
- joins two above frames to distinct beetween "weekend" and "weekday" readings
```{r}
activity_na_imput <- merge(activity_na_imput,weekend_frame, by.x="day",by.y="day")
```
- creates summary dataframe and plots it in two panels to compare "weekend" and "weekday" readings
```{r}
StepsIntervalWeekend <- ddply(activity_na_imput,.(weekend,interval),summarise,mean=mean(steps))
g <- qplot(interval, mean, data = StepsIntervalWeekend)
g <- g + facet_grid(weekend ~.) + geom_line()
g
```
